---
title: "Final Exam"
format: 
  html:
    self-contained: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

Please read all of the questions and tables carefully and follow all instructions. Each question has an allotted point value. Be as thorough and specific as possible; extra calculations and incorrect information, even in the presence of correct information, will result in point deductions. Be sure to show all formulas and all necessary work to answer the questions. You may upload your completed exam to the Elms course site.

::: callout-note
While this is an open-note exam, you are not to receive assistance from anyone else. As usual, the Honor Code applies.
:::

**Total points: 50**

## Conceptual Questions

Please include all work (and computations) necessary to answer the questions.

**Total points: 20 (2 points each)**

### Question 1

Suppose you estimate a bivariate regression model with a total sample size of 50 and obtain a standard error for $\beta_1$ of 2.50. What is the smallest regression coefficient ($\beta_1$) that you could have and still be able to reject the null hypothesis (of no relationship between $X_1$ and $Y$) at the 0.05 level with a one-tailed test?

Sample size: 40

DF: 49

Critical Value one-tailed test 0.05 level: 1.676

B1= 0+1.676\*(2.50) = 4.19

ANSWER: The smallest regression coefficient that you could have and still be able to reject the H0 for a one-tailed, 0.05 significance level is 4.19.

### Question 2

I regress $Y$ on $X_1$ (i.e., $Y = \beta_0 + \beta_1 X_1$) and find that $\beta_1 = 4.20$, $SE_{\beta_1} = 1.90$, $t = 2.21$, $p = 0.01$, and a 95% confidence interval is \[0.48, 7.92\]. What is my best estimate of the effect of a two-unit change in $X$ on the mean of $Y$ in the population?

-   Y = 2\*4.20 = 8.40

-   ANSWER: The best estimate of the effect of a two unit change in X1 on the mean of Y in the population is about 8.40 units

### Question 3

I conduct an OLS regression with a sample size of 90 and 5 independent variables. To determine a p-value for each coefficient, I would examine a t-distribution with how many degrees of freedom?

-   DF = n - 5 - 1 (# of coefficients estimated)

-   DF = 90 - 6 = 84

-   ANSWER: You would examine a t-distribution with 84 degrees of freedom when determining the p-values for each coefficient in your OLS regression

### Question 4

Suppose I regress $Y$ on $X$ and compute the mean response for $Y$ at some specified value of $X$. When determining the confidence interval around this mean response, which of the following will **NOT** have any effect on the width (or, size) of that confidence interval?

A.  Total sample size,

B.  Mean-squared error,

C.  The specified value of $X$,

D.  None of the above: all of these (above) will affect the confidence interval,

E.  There is not enough information to answer this question.

ANSWER: C, because the width of the CI for the mean response in regression analysis depends on the SE of the mean response and since the SE is affected by the mean-squared error, it depends on the total sample size. I also don't use X to calculate the CI.

-   Since the specified value of X can affect the precision of the estimate at that specific point, it doesn't directly affect the CI for the mean response.

### Question 5

I regress $Y$ on $X$ and find that $\beta_1$ has a two-tailed p-value of 0.04. Which of the following statements is most accurate?

A.  The lower and upper bound of a 95% confidence interval around $\beta_1$ will have the same sign,

B.  The absolute value for the t-statistic for $\beta_1$ will be greater than 1.96,

C.  A 90% confidence interval around $\beta_1$ will not contain zero,

D.  All of the above,

E.  None of the above and/or there is not enough information.

-   ANSWER: D

### Question 6

I regress $Y$ on three independent variables -- $X_1$, $X_2$, and $X_3$ -- and I find the following 95% confidence intervals -- $\beta_1$: \[0.12, 1.45\], $\beta_2$: \[-0.01, 0.15\], and $\beta_3$: \[-0.64, -0.01\]. Which of the following statements is most accurate?

A.  $\beta_2$ and $\beta_3$ will have negative coefficients, and $\beta_1$ is statistically significant at the 0.05 level (two-tailed),

B.  $\beta_1$ will have a positive coefficient, and $\beta_1$ is the only statistically significant coefficient (of the three coefficients) at the 0.05 level (two-tailed),

C.  $\beta_1$ will have a positive coefficient, $\beta_2$ and $\beta_3$ will have negative coefficients, and only $\beta_1$ and $\beta_3$ are statistically significant at the 0.05 level (two-tailed),

D.  All of the above,

E.  None of the above and/or there is not enough information.

-   ANSWER: C, because B2 and B3 do not include zero, so they are not statistically significant but B1 does not include zero, leaving it statistically significant. B1 will be positive because it doesn't include zero and will be positive.

### Question 7

Suppose I estimate a regression with two independent variables and obtain a $R^2$ of 0.40 where the Residual Sum of Squares is equal to 5,150. What does the Total Sum of Squares equal in this regression model?

```{r}
#Given values
r_sq <- 0.40
residual_sum_sq <- 5150

# Calculate Total Sum of Squares (SST)
total_sum_sq <- residual_sum_sq / (1 - r_sq)

# Display the result
print(total_sum_sq)

```

-   Total Sum of Squares (SST) in this regression model is 8,583.33.

### Question 8

I regress $Y$ on $X_1$ and $X_2$ (i.e., $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2$). I find that $\beta_1 = -0.87$ with a two-tailed p value of 0.001 and $\beta_2 = 3.52$ with a two-tailed p-value of 0.04. Which of the following statements is most accurate?

A.  $X_2$ has a larger substantive effect on $Y$ than $X_1$, but the effect of $X_1$ is more statistically significant,

B.  $X_1$ has a larger substantive effect on $Y$ than $X_2$ and $X_1$ is more statistically significant than $X_2$,

C.  $X_1$ has a larger substantive effect on $Y$ than $X_2$, but the effect of $X_2$ is more statistically significant,

D.  None of the above and/or there is not enough information.

-   ANSWER: A

### Question 9

I regress $Y$ on $X_1$ (i.e., $Y = \beta_0 + \beta_1 X_1$) and obtain a $R^2$ of 0.45. Then, I regress $Y$ on both $X_1$ and $X_2$ (i.e., $Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2$) and obtain a $R^2$ of 0.53. Which of the following statements must be true?

A.  The correlation between $X_1$ and $Y$ is stronger than the correlation between $X_2$ and $Y$,

B.  The coefficient on $\beta_1$ is statistically significant in both models,

C.  The model with $X_1$ and $X_2$ explains more of the variation in $Y$ than a model with just $X_1$,

D.  All of the above: all of these statements are true.

-   ANSWER: C, because when you have just X1 you get a lower rsq. But when you include X2 you get bigger rsq explaining more of the variation.

### Question 10

Based on the following regression equation, might the negative coefficient on the South variable be the result of southern states having a lower proportion of high school graduates than non-southern states? In no more than one sentence, explain your answer.

$$
Turnout = 34 + 0.5(Percent\ High\ School\ Graduates) – 5.9(South)
$$

-   ANSWER: The negative coefficient does not necessarily imply that southern states have a lower proportion of HS grads than non-southern states, the coefficient only measures the association between the southern states and voter turnout.

## Applied Questions

All datasets referenced below are posted on the ELMs course site.

### Question 1

*Points: 15*

Use the `world` dataset to answer the following questions. The relevant variables and their coding information are as follows: `literacy` indicates a country's literacy rate; `dem_score14` represents a country's level of democratization (higher values indicate greater democratization); `spendeduc` reflects the amount of public expenditures on education as a percentage of GDP; `gdp_10_thou` represents GDP per capita (one unit is \$10,000 USD); `educ_quality` indicates the average quality rating of a country's educational system; and `ungr9095` represents the percent average annual population growth.

#### Part A

*Points: 4*

Evaluate the bivariate relationship (using R) between a country's support for public education and its literacy rate. Report the results in a professionally formatted table. Do countries that spend more money on education appear to have a higher literacy rate?

```{r}
library(readr)
world <- read_csv("~/Desktop/GVPT 622 Lab/world.csv")
View(world)
library(tidyverse)
library(knitr)
library(broom)
library(modelsummary)
library(stargazer)

#regression model 
reg_model1 <- lm(literacy ~ spendeduc, data = world)

summary(reg_model1)

#clean it up
tidy_results1 <- tidy(reg_model1)
tidy_results1

#professional table 
custom_names1 <- c("Education Spending (%)", "Intercept")
stargazer(reg_model1, 
          covariate.labels = custom_names1,
          dep.var.labels = c("Literacy Rate"),
          ci = TRUE,
          ci.level = 0.95,
          type = "text")


```

-   ANSWER: The table suggest that on average, a one unit increase in Education Spending is associated with a 1.722 unit increase in the Literacy Rate.

-   With a positive coefficient, we see a positive relationship between a country support for public education and it's literacy rate.

-   R-sq is 0.027, only 2.7% of the variability in the DV is explained by the IV.

-   It is marginally statistically significant at the .05 significance level.

-   Yes, countries that spend more money on education appear to have a higher literacy rate.

#### Part B

*Points: 5*

Estimate a second regression model (using R) that includes both a country's public expenditures on education and its level of democratization as independent variables. When controlling for democratization, does education expenditures exhibit a significant impact on literacy? Report the results in a professionally formatted table. Be sure to discuss the extent to which the results change, and if so, why they changed.

```{r}
#Fit the regression model
reg_model2 <- lm(literacy ~ spendeduc + dem_score14, data = world)

summary(reg_model2)

#clean it up
tidy_results2 <- tidy(reg_model2)
tidy_results2

# Display the results using stargazer
#professional table 
custom_names2 <- c("Education Spending (%)", "Level of Democracy")
stargazer(reg_model2, 
          covariate.labels = custom_names2,
          dep.var.labels = c("Literacy Rate"),
          ci = TRUE,
          ci.level = 0.95,
          type = "text")

modelsummary(reg_model2,
             coef_rename = c(spendeduc = "Public Spending on Education (%)", 
                             dem_score14 = "Country's level of democratization"),
             statistic = c("t={statistic}", "SE={std.error}", "conf.int", "p.value"))
```

-   ANSWER: There isn't sufficient evidence to suggest that public spending on education impacts literacy in this model.

-   The p-value is .748 which is greater than the .05

This shows us that in this model the public spending on education does not have a statistically impact on literacy when controlling for democratization

#### Part C

*Points: 6*

Now estimate a third model (using R) that includes all of the variables listed above. Report the regression results in a professionally formatted table and interpret each coefficient. When interpreting the impact of democratization using this regression model, present/utilize a visual representation (using R) of the expected change in literacy as a function of democratization.

```{r}
library(marginaleffects)
library(plotly)
#Fit the regression model
reg_model3 <- lm(literacy ~ dem_score14 + spendeduc + gdp_10_thou + educ_quality + ungr9095, data = world)

summary(reg_model3)

#clean it up
tidy_results2 <- tidy(reg_model2)
tidy_results2

#professional table 
modelsummary(reg_model3,
             coef_rename = c( dem_score14 = "Country's level of democratization",
               spendeduc = "Public Spending on Education (%)",
             gdp_10_thou = "GDP per Capita (one-unit $10,000)",
             educ_quality = "Quality of Countrys Education",
             ungr9095 = "Average Annual Population Growth (%)"),
             statistic = c("t={statistic}", "SE={std.error}", "conf.int", "p.value"))

#expected change in literacy regression plot 
plot_predictions(reg_model3, condition = "dem_score14")
```

-   Interpretation of each coefficient

    -   Intercept: 55.91, this shows the expected value of the literacy rate when all IVs are = 0.

    -   Dem_score14: 2.55, a one unit increase in the country's democracy level is associated with an expected increase by 2.55 units. Statistically significant at the 0.05 level

    -   Spendeduc: 0.11, not statistically significant and that suggest that public spending on education does not significantly impact literacy rates.

    -   gpd_1_thou: 2.36, a one unit increase in GDP per capital, literacy is expected to increase by 2.36 units but is not statistically significant.

    -   educ_quality: 3.81, a one unit increase in the quality of education in a given country, literacy is expected to increase by 3.81 units. This is marginally statistically significant at the 0.1 level

    -   ungr9095: -3.33, for a one unit increase the average annual population growth rate, literacy is expected to decrease by 3.33 units and is statistically significant at the 0.001 level.

### Question 2

*Points: 15*

Using the `states` dataset, use R to regress the variable `obama08` (the percentage of a state's vote that President Obama received in the 2008 U.S. presidential election) on the following independent variables: `cig_tax` represents the amount of a state's cigarette tax (in dollars); `college` is the percentage of a state's population that graduated college (0 to 100); `union07` is the percentage of a state's workers that are union members (0 to 100); and `south` is a dichotomous variable coded as `1` if the state is located in the South (`0` otherwise). Report the results in a professionally formatted table. Answer the following questions.

```{r}
library(readr)
states <- read_csv("~/Desktop/GVPT 622 Lab/states.csv")
View(states)
```

#### Part A

*Points: 6*

Interpret the substantive results from this regression model. Be sure to discuss the direction, magnitude, and statistical significance of each slope coefficient. Does the `union07` variably exhibit a substantively significant impact on voting behavior in the 2008 election?

```{r}
# Assuming 'data' is your dataset
reg_model_2 <- lm(obama08 ~ cig_tax + college + union07 + south, data = states)

# Display summary
summary(reg_model_2)

#clean it up
tidy_results_2 <- tidy(reg_model_2)
tidy_results_2

#professional table 
modelsummary(reg_model_2,
             coef_rename = c( cig_tax = "States cigarette tax (USD)",
               college = "State's population that graduated college (%)",
             union07 = "State's workers that are union members (%)",
             south = "Southern States"),
             statistic = c("t={statistic}", "SE={std.error}", "conf.int", "p.value"))

```

Interpretations of Coefficients

Cig_tax: 0.523, the p-value of 0.816 the coefficient is not statistically significant.

-   Positive direction. This suggests that the states cigarette tax does not have significant impact on the percentage of state votes for Obama.

College: 0.965, the coefficient for college is statistically significant (p-value \< 0.001).

-   Positive direction. For every one unit increase in the percentage of a state's population that graduated college, we can expect an increase of 0.965 percentage points in a state's vote for Obama, holding other variables constant.

Union07: 0.751, the coefficient is statistically significant (p-value = 0.003511).

-   Positive direction. For every one unit increase in the percentage of a state's workers that are union members, we can expect an increase of 0.751 percentage points in the state's vote for Obama, holding other variables constant.

-   This suggests that the union membership rate is substantively significant in explaining the variation in the percentage of the state's vote for Obama.

South: 1.392, the coefficient for south is not statistically significant (p-value = 0.593179).

-   Positive direction. This suggests that being located in the South does not have a substantively significant impact on the percentage of a state's vote for Obama.

#### Part B

*Points: 2*

Interpret the intercept coefficient. What does this represent and is it meaningful to interpret this coefficient on its own? Why or why not?

Intercept coefficient: 16.13

-   The intercept represents the estimtated value of the DV when all IV are 0. The estimated percentage of the states vote for Obama when all IVs are 0 is 16.13.

-   In this case, I would say that interpreting the intercept on it's own is not meaningful. It is not meaningful because it would be hard to find a real world example where the workers in the state that are union members is 0. I could see possibly interpreting the intercept alone if a states cigarette tax was zero or if cigarettes were no longer legal and we didn't have a tax on them. From my understanding, every state has a college in it. Whether it be a community college or 4 year college, or those who are residents graduated from another state, the graduation rate from college would not be zero and would not be a realistic. Then for the south IV, if it were equal to 0 that means it's a none southern state, I could see interpreting non-southern states but for a Presidential election, the south is pretty significant. It wouldn't be as meaningful.

-   Overall, no, it is not meaningful to interpret the intercept coefficient on its own.

#### Part C

*Points: 2*

What is the expected mean percentage of Obama's vote total in a Southern state with a cigarette tax of \$2.00, a state population where 30% of people are college graduates, and where 15% of workers in a state are union members?

```{r}
# Coefficients from the regression model
intercept <- 16.1280
cig_tax_coef <- 0.5230
college_coef <- 0.9646
union07_coef <- 0.7514
south_coef <- 1.3928

# Given values for the independent variables
cig_tax <- 2.00
college <- 30
union07 <- 15
south <- 1  # Southern state

# Calculate the expected mean percentage of Obama's vote
expected_percentage <- intercept + cig_tax_coef * cig_tax + college_coef * college + union07_coef * union07 + south_coef * south

# Print the result
cat("Expected mean percentage of Obama's vote:", expected_percentage, "%\n")

```

#### Part D

*Points: 3*

What is a 90% confidence interval around the coefficient for the `south` variable.

```{r}
SE_south <- 2.588
df_south <- 45

CI_level <- 0.90

moe_south <- qt((1 + CI_level) / 2, df_south) * SE_south 

CI_level <- c(south_coef - moe_south, south_coef + moe_south)

cat("90% Confidence Interval for the southern state coefficient:", round (CI_level[1], 3), "to", round(CI_level[2], 3), "\n")
```

#### Part E

*Points: 3*

What is the null hypothesis that the F-test in this regression output is testing?

H0 Null hypothesis: cig_tax= 0, college= 0, union7= 0, south= 0

-   Every IV is equal to 0

Ha the Alternative hypothesis: At least one of these coefficient is not equal to 0.

F-test: 12.19 on 4 and 45 DF

ANSWER: The null hypothesis being tested by the F-test is that all of the coefficient IV in this model are = 0

-   The F-test is used to test if there is a significant linear relationship between the DV (obama08) and all of the IV in the model combined. With a low p-value with the F-stat would result in rejecting the null hypothesis. This would suggest that one of the IV is a predictor of the DV (obama08).
