---
title: "Mid-term Exam #2"
subtitle: "Due date: 10 November"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

Please read all of the questions carefully and follow all instructions. Each question has an allotted point value. Be as thorough and specific as possible; extra calculations and incorrect information, even in the presence of correct information, will result in point deductions. Be sure to show all formulas and all necessary work to answer the questions. You may upload your completed exam to the Elms course site (attach to Midterm Exam #2).

::: callout-note
While this is an open-note exam, you are not to receive assistance from anyone else (as usual, the Honor Code applies).
:::

**Total points: 50 points**

## Conceptual questions

::: callout-note
Please include all work (and computations) necessary to answer the questions.
:::

**Total points: 26**

### Question 1

*2 points*

Suppose I am interested in determining if freshman undergraduates at the University of Maryland spend more hours studying in the average week than sophomore undergraduates. I conduct a study in which I take a simple random sample (SRS) of 1200 freshman students and 1200 sophomore students. I find that the freshman students in my sample study for, on average, 412 minutes per week and the sophomore students in my sample study for, on average, 335 minutes per week. The standard error of the difference is 30. What is a 90% confidence interval for the difference between freshman and sophomore students?

```{r}
# Given values
fresh1 <- 412  #mean study time for freshman students
soph2 <- 335  #mean study time for sophomore students
se <- 30   #standard Error of the difference
confidence_level <- 0.90

# critical value (z) for 90% confidence level
z <- qnorm(1 - (1 - confidence_level) / 2)

#margin of error
margin_of_error <- z * se

#Range of confidence interval
lower_bound <- (fresh1 - soph2) - margin_of_error
upper_bound <- (fresh1 - soph2) + margin_of_error

#print the confidence interval
cat("90% Confidence Interval for the Difference: (", lower_bound, ", ", upper_bound, ")\n")

```

90% CI z-score= 1.645

CI_studytime= (412 - 335) +/- 1.645\*(30)

CI_studytime= 77 +/- 49.35

Upperbound = 77 + 49.35 = 126.35

Lowerbound = 77 - 49.35 = 27.65

ANSWER: The difference between freshman and sophmore students at UMD study times, we can be 90% confidence that the time fall between the range of about 27.65 to 126.35 minutes.

### Question 2

*2 points*

Based on the results of my study described in question 1, can you reject the null hypothesis of no statistically meaningful difference in the study habits of sophomore and freshman students? Why or why not?

```{r}
# Given values
fresh1 <- 412  # Mean study time for freshman students
soph2 <- 335  # Mean study time for sophomore students
se <- 30   # Standard Error of the difference
nfresh <- 1200  # Sample size of freshman students
nsoph <- 1200  # Sample size of sophomore students

# Perform a two-sample t-test
t_stat <- (fresh1 - soph2) / sqrt((se^2/nfresh) + (se^2/nsoph))

# Degrees of freedom
df <- nfresh + nsoph - 2

# Calculate the p-value
p_value <- 2 * pt(abs(t_stat), df, lower.tail = FALSE)

# Define the significance level (alpha)
alpha <- 0.1

# Print the t-statistic and p-value
cat("t-statistic:", t_stat, "\n")
cat("p-value:", p_value, "\n")

#Check if the null hypothesis is rejected or not
if (p_value < alpha) {
  cat("Reject the null hypothesis - There is a statistically meaningful difference in study habits.\n")
} else {
  cat("Fail to reject the null hypothesis - There is no statistically meaningful difference in study habits.\n")
}

```

Null Hypothesis: There is no difference of hours spent studying between freshman and sophomore students at UMD.

m1 = population mean study time for freshmans

m2 = population mean study time for sophomores

-   H0= m1 - m2 = 0

-   ANSWER: We can reject the null hypothesis because the time difference would be 0 minutes, no difference. but the 90% confidence interval of 27.65 to 126.35 minutes, 0 does not fall between that.

    -   There is a statistical difference in the study times between freshman and sophomore students at UMD

### Question 3

*2 points*

If I am testing a null hypothesis that X has no effect on Y in the population (and thus my alternative hypothesis is that X does have an effect), why might I prefer to commit a Type-II error over a Type-I error (and, of course, this holds aside my first preference of making no error at all)? Answer in no more than two sentences.

-   ANSWER: You may prefer to commit a Type II (failing to reject H0 w/ an effect) error over a Type I error (reject the H0 no effect) if the consequence of the error you would commit incorrectly found that X has an effect when it doesn't. This is worse than failing to detect a real effect when its there.

### Question 4

*1 point*

When conducting a difference-of-means test, which of the following samples would yield a sampling distribution with the lowest variability?

A.  A sample of 900 with a standard deviation of 15
B.  A sample of 25 with a standard deviation of 10
C.  A sample of 625 with a standard deviation of 20
D.  A sample of 100 with a standard deviation of 6

ANSWER: A; a larger sample size tends to reduce the variability.

### Question 5

*1 point*

Which of the following probabilities is not independent?

A.  The probability that the roulette wheel will end up on 23 two times in a row.
B.  The probability that three successive coin tosses will each turn up heads.
C.  The probability that I draw an ace and then a king in a row from a deck of cards (when drawing a two-card hand).
D.  The probability that I will get a 6 on three subsequent die rolls.
E.  None of the above -- all are independent probabilities.

ANSWER: C, the probabilty that I will draw an ace, then a king in a row from a deck of cards is not independent. The probability of drawing a king depends on if I've already drawn an ace. These are not independent of eachother.

### Question 6

*1 point*

I conduct a two-tailed difference-of-means test and obtain a t-statistic of 2.10. Is my result statistically significant (with 95% confidence)?

A.  Yes, at the 0.05 level (p\<0.05)
B.  Yes, but only at the 0.10 level (p\<0.10)
C.  Not at either the 0.05 or the 0.10 level
D.  There is not enough information to answer this question.

ANSWER: D; There is not enough information to answer this question, becauase I need to know the degrees of freedom or the critical value

### Question 7

*1 point*

I take a sample of 1800 adults and find that 360 of them watched last Monday's NFL game. What probability represents the complement to the sample proportion of adults who watched the NFL game?

-   Complements must sum to 1

-   Sample proportion: 360/1800= .2

-   1-.2= 0.8

ANSWER: The complement probability (that an adult did not watch the game), is 0.8

### Question 8

*1 point*

Which of the following makes it more likely that a given sample statistic will be statistically different from zero (and thus allow you to reject the null hypothesis, all else equal)?

A.  More observations
B.  Greater variance in the sample
C.  Using a two-tailed instead of a one-tailed significance test
D.  A larger confidence interval around the test statistic
E.  Both (a) and (b) (but not (c) or (d))
F.  All of the above increase the chances of statistical significance
G.  None of the above

ANSWER: E; Both (a) and (b) (but not (c) or (d)).

Because:

C: this would increase the requirements of the test and actually make it harder to reject the H0 compared to a one-tailed test

D: a larger CI indicates a larger uncertainty about the estimate given and that would be the opposite of trying to reject the H0

### Question 9

*1 point*

The p-value for a two-tailed test of the null hypothesis $H_0: \mu=40$ is 0.06. Which of the following statements is accurate?

A.  A 95% confidence interval around the sample mean includes the value 40
B.  A 90% confidence interval around the sample mean includes the value 40
C.  A 99% confidence interval around the sample mean does not include the value 40
D.  None of the above statements are correct
E.  All of the above statements are correct

ANSWER: B; A 90% confidence interval around the sample mean includes the value 40

-   The p-value of .06 for a two tailed test means that if the H0 is true )population mean is 40), I can expect to see you mean sample about 6% of the time., at the 95% we would see it 5% of the time. This is not within our confidence interval for 95%, therefore the answer is B.

### Question 10

*4 points total*

#### Part A

*2 points*

If you roll a fair (six-sided) die twice, what is the probability that both rolls will be odd or greater than four?

-   P(ODD) Probability of rolling an odd number= 1,3,5; 3/6

-   P(G) Probability of rolling a number \> 4= 5,6; 2/6, 1/3

-   Probability that I get an odd on the first and second roll: 1/2 \* 1/2 = 1/4

-   Probability that I get number \> 4 = 1/3 \* 1/3 = 1/9

-   1/9 + 1/4 = 13/36 = 13/36

ANSWER: The probability that both rolls will be odd or greater than four = 13/36, 0.36

#### Part B

*2 points*

The following is a distribution of U.S. college students classified by their age and full- vs. part-time status. Based on these data, what is the probability that a student is in the 25-29 age group, given that (i.e., conditional on knowledge that) the student is full time?

```{r}
library(tidyverse)

tibble::tibble(
  age = c("15 - 19", "20 - 24", "25 - 29", "30+"),
  full_time = c(155, 255, 75, 35),
  part_time = c(20, 55, 80, 95)
) |> 
  knitr::kable()

```

Find probability that a given student is full time, but in the age range of 25-29

```{r}

data <- tibble(
  age = c("15 - 19", "20 - 24", "25 - 29", "30+"),
  full_time = c(155, 255, 75, 35),
  part_time = c(20, 55, 80, 95)
)

#the conditional probability
age_group <- "25 - 29"
condition_full_time <- data$age == age_group
probability <- sum(data$full_time[condition_full_time]) / sum(data$full_time)

#print the result
cat("The probability that a student is in the", age_group, "age group, given that the student is full-time, is:", probability, "\n")

```

ANSWER: the probability that a student is in the 25-29 age group, given that (i.e., conditional on knowledge that) the student is full time = 0.144 or 75/520

### Question 11

*4 points total*

#### Part A

*2 points*

Using a SRS of 1211 people, I estimate that the proportion of people living in the South that favor teaching sexual education in public schools is 0.874 and the proportion of people outside of the south that favor teaching sexual education in public schools is 0.906. And, the standard error of the difference (in citizen views about teaching sexual education in public schools) between people living in the south and those not living in the south is 0.015. Give an interval estimate for the difference in the proportion of people favoring sex education in public schools between people who do, and do not, live in the south.

```{r}
#the given values
pe <- 0.874 - 0.906  #point Estimate
se <- 0.015  #standard error
alpha <- 0.05  #significance level 95% CI

#critical value for a two-tailed test 95% CI
z <- qnorm(1 - alpha / 2)

#margin of error
moe <- z * se

# Calculate the confidence interval
lower_Bound <- pe - moe
upper_Bound <- pe + moe

# Print the confidence interval
cat("95% Confidence Interval for the difference in proportions:", lower_Bound, "to", upper_Bound, "\n")


# Given values
p1 <- 0.874  # Proportion in the South
p2 <- 0.906  # Proportion outside the South
SE <- 0.015  # Standard error of the difference
Z <- 1.96    # Z-score for 95% confidence interval

# Calculate the confidence interval
interval <- c((p1 - p2) - Z * SE, (p1 - p2) + Z * SE)

# Print the result
print(interval)

```

lowerbound: (0.874 - 0.906) - (1.96 \* 0.015) = -0.0614

upperbound: (0.874 - 0.906) + (1.96 \* 0.015) = --0.0026

ANSWER: The interval estimate for the difference in the proportion of people favoring sex education in public schools between people who do, and do not, live in the south is -0.0614 to -0.0026.

#### Part B

*2 points*

Do the data (i.e., estimates above) show support for my hypothesis? How do you know?

```{r}
#given values to test hypothesis
pr_in_south <- 0.874
pr_not_south <- 0.906
se <- 0.015
n <- 1211  
alpha <- 0.05  

# Calculate the test statistic (Z-score)
t_stat <- (pr_in_south - pr_not_south) / se

# Find the critical values for a two-tailed test at the specified alpha level
c_value <- qnorm(1 - alpha / 2)

# Calculate the p-value for the test
p_value <- 2 * (1 - pnorm(abs(t_stat)))

# Print the test statistic, critical value, and p-value
cat("Test Statistic:", t_stat, "\n")
cat("Critical Value:", c_value, "\n")
cat("P-Value:", p_value, "\n")

# Check whether to reject the null hypothesis
if (abs(t_stat) > c_value) {
  cat("Reject the null hypothesis - The proportions are statistically different.\n")
} else {
  cat("Fail to reject the null hypothesis - The proportions are not statistically different.\n")
}

```

ANSWER: The t statistic is outside of the critical value and p-value is less than our 0.05 confidence level. I will reject the H0 and the data provides evidence in support for the hypothesis.

### Question 12

*6 points total*

I am interested in estimating the average number of texts that University of Maryland undergraduate students send in a day. My hypothesis is that the average number of texts students send is greater than 100. I take a SRS of 1600 students and find that the mean number of texts they send is 105, and with a standard deviation of 120.

#### Part A

*2 points*

What is a 95% confidence interval around the sample statistic?

```{r}
#given values
sample_mean <- 105
pop_sd <- 120
sample_size <- 1600
CI_level <- 0.95

#critical value for a 95% CI
c_value2 <- qnorm(1 - (1 - CI_level) / 2)

#calculate the margin of error
moe_2 <- c_value2 * (pop_sd / sqrt(sample_size))

#calculate the confidence interval
lower_bound2 <- sample_mean - moe_2
upper_bound2 <- sample_mean + moe_2

#print the confidence interval
cat("95% Confidence Interval: [", lower_bound2, ", ", upper_bound2, "]\n")

```

ANSWER: 95% confidence interval around the sample statistic is 99.12 to 110.88

#### Part B

*2 points*

When testing the null hypothesis, what is the test statistic associated with the sample statistic?

$$
t=105-100/(120)/\sqrt1600 = 5/3, 1.67
$$\

-   The test statistic associated with the sample statistic is about 1.67

-   The test statistic allows me to test the null for a one sample t-test and that's the equation I used above.

#### Part C

*2 points*

If using a one-tailed test of the null hypothesis and you are willing to accept a Type-I error rate of 0.05, do the data support my hypothesis? Why or why not?

```{r}
library(stats)
p_value <- 1 - pt(1.67, 1599)
p_value
```

ANSWER: Critical value for one tailed test 0.05 significance level 1.645, and our t-value is 1.67. My p-value is less than confidence level .05, I would reject the null-hypothesis and the data supports the hypothesis.

-   I would accept the Type-I error rate of 0.05 and the data support the hypothesis

## Applied questions

::: callout-note
All datasets referenced below are available through the course elms page. Do note use the `poliscidata` package for these questions, as the specific variables referenced are not all included in that package.
:::

**Total points: 24**

### Question 1

*8 points total*

I hypothesize that, among only those that were eligible to vote, people with greater confidence in the U.S. military were more likely to turnout to vote in the 2012 presidential election. Use data from the General Social Survey (i.e., the `gss` dataset) to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `conarmy` (1 = a "great deal" of confidence; 2 = "only some" confidence; 3 = "hardly any"); and `vote12` (1 = voted; 2 = did not vote; 3 = ineligible). Answer the following questions.

#### Part A

*3 points*

Complete a cross-tab and interpret the results. Do the data support my hypothesis? Be sure to explain the nature of the relationship (or lack thereof, if relevant).

```{r}
library(tidyverse)
library(dplyr)
# Load the 'gss' dataset

library(readr)
gss <- read.csv("gss.csv")

gss_data <- gss |>
  select(vote12, conarmy) |>
  drop_na()

```

```{r}
library(modelsummary)
datasummary_crosstab(vote12 ~ conarmy, data = gss_data)
```

#### Part B

*3 points*

Compute (by hand) the chi-square statistic to test the null hypothesis of no relationship between these two variables. Be sure to show your work.

$$
\chi^2 = \Sigma{\frac{(Observed\ count - Expected\ count)^2}{Expected\ count}}
$$

$$
Expected Count = \frac{Row\ total * Column\ total}{N}
$$

Confidence Level in Military

+------------------------+----------------------+----------------------+---------------------+--------------+
| Confidence in Military | Great deal (1)       | Only Some (2)        | Hardly (3)          | Total        |
+========================+======================+======================+=====================+==============+
| Voted (1)              | Observed: 642, 52.8% | Observed: 498, 41%   | Observed: 75, 6.2%  | 1215         |
|                        |                      |                      |                     |              |
|                        | Expected: 638.9      | Expected: 483.82     | Expected: 92.28     | Chi: 3.67    |
|                        |                      |                      |                     |              |
|                        | Chi: 0.02            | Chi:0.42             | Chi:3.24            |              |
+------------------------+----------------------+----------------------+---------------------+--------------+
| Did Not Vote (2)       | Observed: 278, 51.8% | Observed: 208, 38.7% | Observed: 51, 9.5%  | 537          |
|                        |                      |                      |                     |              |
|                        | Expected: 282.38     | Expected: 213.84     | Expected: 40.78     | Chi: 2.279   |
|                        |                      |                      |                     |              |
|                        | Chi:0.07             | Chi:0.16             | Chi: 2.56           |              |
+------------------------+----------------------+----------------------+---------------------+--------------+
| Ineligible (3)         | Observed: 77, 53.5%  | Observed: 49, 34%    | Observed: 18, 12.5% | 144          |
|                        |                      |                      |                     |              |
|                        | Expected: 75.72      | Expected: 57.34      | Expected: 10.94     | Chi: 5.79    |
|                        |                      |                      |                     |              |
|                        | Chi:0.02             | Chi: 1.21            | Chi: 4.56           |              |
+------------------------+----------------------+----------------------+---------------------+--------------+
| Total                  | 997                  | 755                  | 144                 | 1896         |
|                        |                      |                      |                     |              |
|                        | 52.6%                | 39.8%                | 7.6%                | Chisq: 12.24 |
+------------------------+----------------------+----------------------+---------------------+--------------+

Chi squared: = 12.24

#### Part C

*2 points*

Using the chi-square statistic that you computed in question 1(b), can you reject the null hypothesis of no relationship between these two variables with 95% confidence? Why, or why not?

ANSWER: I can reject the null of no relationship between these two variables with 95% confidence because my chisq statistic = 12.24, df = 4 and the critical value for rejection is 9.49 at the 95% confidence level. The chisq statistic is greater than the critical value, which rejects the null hypothesis at 95% CI

### Question 2

*8 points total*

I hypothesize that citizens who do not support fracking are less conservative than those who do support fracking. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `fracking` (1 = "approve" of fracking; 2 = "middle"; 3 = "disapprove"); and `libcon7` (higher values represent less liberalism, or more conservatism). Answer the following questions.

```{r}
library(tidyverse)
library(modelsummary)
library(dplyr)

library(readr)
nes_data <- read_csv("nes.csv")

nes_data <- nes_data |>
  select(fracking, libcon7) |>
  drop_na()

nes_data

datasummary_crosstab(libcon7 ~ fracking, data = nes_data)
```

#### Part A

*2 points*

Using these data, what is the point estimate for the mean conservatism/liberalism score among those that disapprove of fracking? What is the point estimate for the mean conservatism/liberalism score among those that approve of fracking?

```{r}
library(dplyr)

#score among disapprove of fracking 
m_disapprove <- nes_data |>
  filter(fracking == 3) |>
  summarise(m_libcon7_disapprove = mean(libcon7))

#score among approve of fracking
m_approve <- nes_data |>
  filter(fracking == 1) |>
  summarise(m_libcon7_approve = mean(libcon7))

#results
m_disapprove
m_approve
```

ANSWER: The point estimate for the mean conservatism/liberalism score among those that disapprove of fracking is 3.55.

-   The point estimate for the mean conservatism/liberalism score among those that approve of fracking is 5.11

#### Part B

*6 points*

Evaluate the null hypothesis that there is no difference in the mean conservatism/liberalism score among those that approve vs. disapprove of fracking. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).

```{r}
#give values to find t and p values 
m_approve <- 5.109453	
m_disapprove <-3.54943
diff_mean <- m_disapprove-m_approve
diff_mean

#standard deviation for approve and disapprove 
stdev_disapprove <- nes_data |>
  filter(fracking == 3) |>
  summarise(stdev_libcon7_disapprove = sd(libcon7))
stdev_disapprove 	

stdev_approve <- nes_data |>
  filter(fracking == 1) |>
  summarise(stdev_libcon7_approve = sd(libcon7))
stdev_approve 

#sample size both approve and disapprove 
sample_approve <- nes_data |>
  filter(fracking == 1) |>
  summarise(sample_libcon7_approve = length(na.omit(libcon7)))

sample_disapprove <- nes_data |>
  filter(fracking == 3) |>
  summarise(sample_libcon7_disapprove = length(na.omit(libcon7)))

sample_approve <- 804
sample_disapprove <-1578

#Standard deviation together - found online equation 
sd_both <- sqrt(((1578-1)*(1.49118^2)+(804-1)*(1.274507^2))/((1578+804)-2))
sd_both

#t-stat
t_stat_q2b <-diff_mean/(1.421771)*(sqrt(1/1578+1/804))
t_stat_q2b

#df
df_q2b <- (804+1578)-2
df_q2b

#two-tailed p-value
p_value_q2b <- 2 * (1 - pt(abs(-0.04754345), 2380))

p_value_q2b
```

ANSWER: since the p-value is greater than 0.05, showing that there is not enough evidence to reject the H0. There is no statistical difference in the mean conservatism/liberalism score of approving and disapproving of fracking.

### Question 3

*8 points total*

I hypothesize that people who express that religion is important to them were more likely to turnout to vote in the 2016 presidential election. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `Relig_imp` (0 = not important; 1 = somewhat important; 2 = quite a bit; 3 = a great deal); and `Voted_2016` (0 = did not vote; 1= voted). Answer the following questions.

#### Part A

*2 points*

Using these data, what is the point estimate for the proportion of respondents that voted (i.e., turnout rate), among citizens expressing that religion is not important? What is the point estimate for the proportion of respondents that voted, among citizens expressing that religion matters a great deal?

```{r}
nes_data <- read.csv("nes.csv")
nes_data2 <- nes_data |>
  select(Relig_imp, Voted_2016) |>
  drop_na()

datasummary_crosstab(Voted_2016 ~ Relig_imp, data = nes_data2)
```

```{r}
v_not_imp_relig <- 971
total_not_important_n <- 1268

pe_v_not_relig <-v_not_imp_relig/total_not_important_n

pe_v_not_relig
```

ANSWER: The point estimate for the proportion of respondents that voted (i.e., turnout rate), among citizens expressing that religion is not important = 0.766

```{r}
v_gd_relig <- 995
total_gd_n <- 1229

pe_v_gd_relig <-v_gd_relig/total_gd_n

pe_v_gd_relig
```

ANSWER: The point estimate for the proportion of respondents that voted, among citizens expressing that religion matters a great deal = 0.810

#### Part B

*6 points*

Evaluate the null hypothesis that there is no difference in the proportion of voters (i.e., turnout rate) among citizens expressing that religion is not important vs. those reporting that religion matters a great deal. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).

+-------+----------------------+----------------------+----------------------+----------------------+--------------+
| Voted | 0                    | 1                    | 2                    | 3                    | All          |
+=======+======================+======================+======================+======================+==============+
| 0     | Observed: 297, 38.2% | Observed: 106, 13.6% | Observed: 141, 18.1% | Observed: 234, 30.1% | 778          |
|       |                      |                      |                      |                      |              |
|       | Expected: 272.51     | Expected: 101.655    | Expected: 139.696    | Expected: 264.133    | Chi: 5.836   |
|       |                      |                      |                      |                      |              |
|       | Chi: 2.201           | Chi: 0.186           | Chi: 0.012           | Chi: 3.438           |              |
+-------+----------------------+----------------------+----------------------+----------------------+--------------+
| 1     | Observed: 971, 34.2% | Observed: 367, 12.9% | Observed: 509, 17.9% | Observed: 995, 35%   | 2842         |
|       |                      |                      |                      |                      |              |
|       | Expected: 995.485    | Expected: 371.344    | Expected: 510.304    | Expected: 964.866    | Chi: 1.597   |
|       |                      |                      |                      |                      |              |
|       | Chi: 0.602           | Chi: 0.051           | Chi: 0.003           | Chi: 0.941           |              |
+-------+----------------------+----------------------+----------------------+----------------------+--------------+
| All   | 1268                 | 473                  | 650                  | 1229                 | 3620         |
|       |                      |                      |                      |                      |              |
|       |                      |                      |                      |                      | Chisq: 7.433 |
+-------+----------------------+----------------------+----------------------+----------------------+--------------+

```{r}
# Degrees of freedom
df <- 3

# Significance level (alpha)
alpha <- 0.05  # For example, a significance level of 0.05 (5%)

# Finding critical value
critical_value <- 7.814
critical_value
```

-   chisq stat = 7.433

-   df = 3

-   critical value for rejection = 7.81

ANSWER: The calculated chisq stat is 7.43 with df = 3 and the critical value of rejection is 7.81 at the 95% CI.

-   The data does not support your hypothesis

-   My calculated chisq stat is less than my critical value of rejection, this suggests that the data fails to reject the null hypothesis and that there is a difference in proportion between citizens citizens that express religion-is-not-important vs. voters that express that religion matters-a-great-deal.
