---
title: "Problem Set 1"
subtitle: "Due date: 18 September"
format: 
  html:
    self-contained: true
editor: visual
execute: 
  echo: true
---

```{r}

library(tidyverse)
library(poliscidata)
```

Please upload your completed assignment to the ELMs course site (under the assignments menu). Remember to include an annotated script file for all work with R and show your math for all other problems (if applicable, or necessary). Please also upload your completed assignment to the Github repository that you have shared with us. *We should be able to run your script with no errors.*

**Total points: 25**

::: {.callout-tip title="HG comments"}
Great work and code. Just please be careful to answer all of the questions in full. 21/25.
:::

## Question 1

*Points: 5*

::: {.callout-tip title="HG comments"}
Good answer and code. Need more detail about the distributions and I am missing your frequency distribution in the rendered document. 3/5.
:::

Using the `gss` dataset (which has survey data with the individual respondent as the unit of analysis), create a frequency distribution and bar chart for each of the following variables: `degree` and `partyid_3`. Describe the distribution of each variable in detail.

::: callout-note
The `gss` dataset can be found in `poliscidata::gss`. Take a look at `?gss` to see what these variables measure.
:::

```{r}
degree_data <- select(poliscidata::gss, degree)

degree <- table(degree_data)

barplot(degree, main = "Frequency of Degree", xlab = "Degree", ylab = "Frequency")

summary(degree)
```

::: {.callout-tip title="HG comments"}
Please make sure to print out the frequency distribution so that we can see it. I cannot see it in your rendered document (you need to explicitly call it after you have assigned it to an object to print it). 

For example: 

```{r}
degree_data <- select(poliscidata::gss, degree)

degree <- table(degree_data)
degree
```
:::

ANSWER: The highest number in the distribution is from people with HS diplomas

::: {.callout-tip title="HG comments"}
Because the question asked for you to describe the distribution of each variable in detail, we need more detail about the distribution than just the most frequently occurring category. 
:::

```{r}
partyid_data <- select(poliscidata::gss, partyid_3)

partyid <- table(partyid_data)

barplot(partyid, main = "Frequency of partyid", xlab = "partyid", ylab = "Frequency")

summary(partyid)
```

ANSWER: The distribution describes the party ID of voters, in this frequency table, the independents have the highest number.

::: {.callout-tip title="HG comments"}
Please see comments above. 
:::

## Question 2

Examine the following vector of (fake) student IQ test scores:

```{r}
iq <- c(
  145, 139, 126, 122, 125, 130, 96, 110, 118, 118, 101, 142, 134, 124, 112, 109, 
  134, 113, 81, 113, 123, 94, 100, 136, 109, 131, 117, 110, 127, 124, 106, 124, 
  115, 133, 116, 102, 127, 117, 109, 137, 117, 90, 103, 114, 139, 101, 122, 105, 
  97, 89, 102, 108, 110, 128, 114, 112, 114, 102, 82, 101
)
```

### Part A

*Points: 5*

::: {.callout-tip title="HG comments"}
Correct and good code. 5/5.
:::

Find the five-number summary, mean, and standard deviation for these data. Also, are there any suspected outliers in the distribution? If so, what are they and how do you know?

```{r}
summary(iq)
iq_sd <- sd(iq)
iq_sd

mean(iq)
sd(iq)

iq.df <- data.frame(iq)

ggplot(iq.df, aes(x = "", y = iq)) +
  geom_boxplot(fill = "black") +
  labs(title = "iq boxplot", y = "Student Scores") +
  theme_minimal()

boxplot.stats(iq)$out

outlier <- boxplot.stats(iq)$out
```

::: {.callout-tip title="HG comments"}
I recommend picking a light fill colour. The solid black makes it difficult to see the median line. 
:::

ANSWER: SD: 14.80 MEAN: 114.98

ANSWER: According to the presented the box plot there seems to be no *suspected* outliers.

### Part B

*Points: 5*

::: {.callout-tip title="HG comments"}
Missing your comparison of the SDs. 3/5.
:::

In large populations, IQ scores are standardized to have a mean of 100 and a standard deviation of 15. In what way does the distribution among these students differ from the overall population?

```{r}
mean <- mean(iq)

sd <- sd(iq)

standard <- (iq - mean)/sd * 15 + 100 

barplot(iq, main = "IQ scores", xlab = "student scores", ylab = "mean") 


```

::: {.callout-tip title="HG comments"}
A bar plot is not the best graph type to use to detect distributions. This is because the order on the x-axis is arbitrary. A histogram or density plot are better graph types for exploring this question. These graph types provide a more meaningful order to our data, allowing us to get a better sense of their central point and spread. 

```{r}
ggplot(tibble(iq = iq), aes(x = iq)) + 
  geom_density()
```
:::

ANSWER: The distribution in this barplot is a bit hard to see but I believe that the distribution with these students differ from the overall population because their IQ is a bit higher. Testing/scores can be dependent on environmental conditions that are not taken into account (noise, test anxiety, sleep etc).

::: {.callout-tip title="HG comments"}
Correct: the mean IQ score for this group is higher than the mean found (by design) within large populations. 

We are missing your assessment of the difference between this group's standard deviation and that of a large population. The standard deviation of this group is `r round(sd(iq), 2)`. The standard deviation of large populations is 15. Therefore, these two groups show roughly the same spread. This group is slightly more clustered around the mean (has a smaller variance) than that of a large population. 
:::

## Question 3

*Points: 5*

::: {.callout-tip title="HG comments"}
Correct. 5/5.
:::

A polling firm is interested in determining how different characteristics of an individual affect vote choice. They record the following characteristics for each survey respondent - race, age, household income, and party affiliation. What "type" of variable does each characteristic most likely represent?

ANSWER: Race is the categorical variable

Age is the continuous variable

Household income is the continuous

Party affiliation is the Categorical

## Question 4

*Points: 5*

::: {.callout-tip title="HG comments"}
Correct. 5/5.
:::

Using the `states` dataset (the unit of analysis is each U.S. state), generate a box plot and density curve for each of the `vep12_turnout` and `cig_tax12` variables. Describe the distribution of each variable in detail.

::: callout-note
The `states` data set can be found in `poliscidata::states`. Take a look at `?states` to see what these variables measure.
:::

::: {.callout-tip title="HG comments"}
Make sure to check your rendered document before submitting. You have printed out the whole data set. No worries! It just makes the document hard to read. I would remove this line of code: `poliscidata::states`.  
:::

```{r}
# poliscidata::states
boxplot(states$vep12_turnout)
boxplot(states$cig_tax12)
plot(density(states$cig_tax12))
plot(density(states$vep12_turnout))

?states
```

Percent turnout of voting eligible population (2012)

-   The density curve shows a normal distribution

Cigarette tax per pack (2012)

-   The density curve shows a skewed right distribution
