---
title: "Problem Set 6"
subtitle: "Due date: 30 October"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
---

Please upload your completed assignment to the ELMs course site (under the assignments menu). Remember to include an annotated script file for all work with R and show your math for all other problems (if applicable, or necessary). Please also upload your completed assignment to the Github repository that you have shared with us. *We should be able to run your script with no errors.*

**Total points: 40**

## Question 1

*Points: 10*

I hypothesize that the mean mathematics SAT score for University of Maryland students is different than 600. I take an SRS of 625 students and find that the mean mathematics score among those students is 610. The standard deviation of my sample is 75.

### Part A

Based on my sample, what is a point estimate for the mean mathematics SAT score for University of Maryland students?

Answer: Based on my sample, the point estimate for the mean is the sample mean. And my sample mean is 610, based on my sample of 625

```{r}
# Define the sample mean
sample_mean <- 610

# Print the point estimate
cat("Point estimate for the mean mathematics SAT score:", sample_mean, "\n")

```

### Part B

What is a 95% confidence interval around that estimate?

```{r}
#sample mean, sample size, and sample standard deviation
sample_mean <- 610
sample_size <- 625
sample_sd <- 75

#onfidence level (95%)
CI95_UMD <- 0.95

#standard error
se <- sample_sd / sqrt(sample_size)

#margin of error
margin_of_error <- qt((1 + CI95_UMD) / 2, df = sample_size - 1) * se

#lower and upper bounds of the confidence interval
lower_bound <- sample_mean - margin_of_error
upper_bound <- sample_mean + margin_of_error

#result
cat("95% Confidence Interval for the Mean Mathematics SAT Score:", 
    round(lower_bound, 2), "to", round(upper_bound, 2), "\n")

```

ANSWER: The confidence Interval of 95% is 604.11 to 615.89

### Part C

What is the null hypothesis?

Null Hypothesis (H0):

The mean mathematics SAT score for University of Maryland students is = 600.

### Part D

What is the two-tailed p-value and what does it mean?

-   The two tailed p-value measures the strength of the evidence against the null hypothesis in a hypothesis test. The test looks for the possibility of a difference in each direction from the null hypothesis

```{r}
# Define the sample mean, sample size, and sample standard deviation
sample_mean <- 610
sample_size <- 625
sample_sd <- 75

# Define the population mean under the null hypothesis
population_mean <- 600

# Calculate the standard error
se <- sample_sd / sqrt(sample_size)

# Calculate the t-statistic
t_stat <- (sample_mean - population_mean) / se

# Degrees of freedom
df <- sample_size - 1

# Calculate the two-tailed p-value
p_value <- 2 * pt(-abs(t_stat), df)

# Print the two-tailed p-value
cat("Two-tailed p-value:", p_value, "\n")

```

### Part E

Do the data support my hypothesis? Why, or why not?

-   The p-value is very small, so I fail to reject the null hypothesis. So the hypothesis presented is not supported by the data.

-   two tailed test p-value is .000908 with a CI level of 95% means that it would very unlikely to observe a sample mean that is 610

## Question 2

*Points: 10*

I take an SRS of 900 citizens in a public opinion survey and find that the mean feeling thermometer rating (on a 0 to 100 scale) of President Biden is 51.5 with a standard deviation of 25. Now, suppose I hypothesize that feelings toward Biden should be greater than 50.

### Part A

What is the null hypothesis?

-   H0: The feeling thermometer rating of President Biden, the mean is equal to 50

### Part B

What is a 95% confidence interval around the sample statistic? Does this interval indicate support for a two-tailed test of my hypothesis (`p < 0.05` as the threshold)? Why, or why not?

```{r}
# Define the sample mean, sample size, and sample standard deviation
sample_mean <- 51.5
sample_size <- 900
sample_sd <- 25

# Set the confidence level (95%)
CI95_Pres <- 0.95

#Calculate the standard error
se <- sample_sd / sqrt(sample_size)

#z-score for a 95% confidence interval
critical_value <- qnorm((1 + CI95_Pres) / 2)

#Calculate the margin of error
margin_of_error <- critical_value * se

#Calculate the lower and upper bounds of the confidence interval
lower_bound <- sample_mean - margin_of_error
upper_bound <- sample_mean + margin_of_error

#Print the confidence interval
cat("95% Confidence Interval for the Mean Rating of President Biden:", 
    round(lower_bound, 2), "to", round(upper_bound, 2), "\n")

```

ANSWER: CI 95% is 49.9 to 53.1

-   This CI support the hypothesis that feelings toward President Biden are greater than 50, I can be reasonably confidence that the true population mean rating falls between this interval

-   The interval does not indicate support for a two tailed test because we cannot reject the null hypothesis = 50, and our CI of 95% range of 49.9-51.1

### Part C

What is a 90% confidence interval around the sample statistic? Does this interval indicate support for a one-tailed test of my hypothesis (`p < 0.05` as the threshold)? Why, or why not?

```{r}
# Define the sample mean, sample size, and sample standard deviation
sample_mean <- 51.5
sample_size <- 900
sample_sd <- 25

#confidence level (90%)
CI90_Pres <- 0.90

#Calculate the standard error
se <- sample_sd / sqrt(sample_size)

#z-score for a 90% confidence interval
critical_value <- qnorm((1 + CI90_Pres) / 2)

#Calculate the margin of error
margin_of_error <- critical_value * se

#Calculate the lower and upper bounds of the confidence interval
lower_bound <- sample_mean - margin_of_error
upper_bound <- sample_mean + margin_of_error

#Print the confidence interval
cat("90% Confidence Interval for the Mean Rating of President Biden:", 
    round(lower_bound, 2), "to", round(upper_bound, 2), "\n")

```

-   CI 90% interval ranges 50.13 to 52.87

-   Since my lower bound is greater than 50 (50.13), it provides support for a one tailed hypothesis, meaning I can be 90% confident that the true mean rating is greater than 50.

## Question 3

*Points: 5*

Suppose I conduct a difference of means test and obtain a t-statistic of 2.50. What does this indicate about the statistical significance of the sample mean compared to the null hypothesis?

-   The t test is an indicator of the difference between the sample mean and the null hypothesis in terms of standard error, how many standard errors away we are from the null hypothesis mean

-   With the information given, I can't indicate anything about the statistical significance of the sample mean compared to the null hypothesis, other than it being 2.5 standard errors away from the null hypothesis

## Question 4

*Points: 5*

I hypothesize that fewer than 40% of registered voters approve of President Biden. I conduct an SRS of 720 registered voters and find that 37.9% of them approve of President Biden.

### Part A

Based on my sample, what is a point estimate for the proportion of the population that approves of President Biden?\

-   The point estimate for the proportion of the population that approves President Biden, is 37.9%

-   based on my sample, the sample proportion observed in the survey. Since we found the 37.9% of the 720 registered voters in the sample approve of Biden

### Part B

What is a 90% confidence interval around that estimate?

```{r}
#sample proportion, sample size, and z-score
sample_proportion <- 0.379
sample_size <- 720
CI90_Pres2 <- 0.90
z_score <- qnorm((1 + CI90_Pres2) / 2)

#Calculate the standard error
standard_error <- sqrt((sample_proportion * (1 - sample_proportion)) / sample_size)

#Calculate the margin of error
margin_of_error <- z_score * standard_error

#lower and upper bounds of the confidence interval
lower_bound <- sample_proportion - margin_of_error
upper_bound <- sample_proportion + margin_of_error

#Print the confidence interval
cat("90% Confidence Interval for the Proportion Approving of President Biden:", 
    round(lower_bound, 4), "to", round(upper_bound, 4), "\n")

```

-   90% CI for the proportions approving of President Biden is 0.349. to 0.408

### Part C

What is the null hypothesis?

-   The null hypothesis = 40%

### Part D

What is the one-tailed p-value and what does it mean?

```{r}
#sample proportion, sample size, and null hypothesis proportion
sample_proportion <- 0.379
sample_size <- 720
null_proportion <- 0.40

#test statistic (z)
z <- (sample_proportion - null_proportion) / sqrt(null_proportion * (1 - null_proportion) / sample_size)

#one-tailed p-value for the lower tail
p_value <- pnorm(z, lower.tail = TRUE)

#one-tailed p-value
cat("One-tailed p-value:", p_value, "\n")

```

-   One-tailed p-value = 0.125

-   This means that based on my sample data, I would fail to reject the null hypothesis. I do not have sufficient evidence to conclude that the proportion of voters who approve of President Biden is less than 40%

### Part E

Do the data support my hypothesis? Why, or why not?

-   The data does not support the hypothesis that fewer than 40% of registered voters approve of President Biden.

-   The one tailed p value 0.125 at a 90% confidence level, the p-value is not small enough to reject the null hypothesis. I do not have enough statistical evidence to support the hypothesis.

## Question 5

*Points: 5*

I hypothesize that in countries where muslim is the predominate religious group, women have a greater number of children on average (i.e., the total fertility rate is higher). Use the `world` dataset to test my hypothesis. State the null hypothesis and interpret what the results tell us about the null and alternative hypotheses. Be sure to show all work necessary to find the answer (i.e., you may use R to assist you, but you should show the necessary computations by hand).

::: callout-note
The `world` data set can be found in `poliscidata::world`.
:::

```{r}
library(tidyverse)
library(dplyr)
poliscidata::world

world <- poliscidata::world|>
  mutate(group_religion = ifelse(religoin == "Muslim", "Muslim", "non_Muslim"))

world_religion_m <- world|>
  filter(group_religion%in%c("Muslim")) |>
  select(group_religion, fertility)

world_religion_n_m <- world |>
  filter(group_religion%in%c("non_Muslim")) |>
  select(group_religion, fertility)

#Test Hypothesis 

#define values
p_hat_5 <- 46/167
p_0_5 <- 119/167
sample_size_religion <- 167

z_t_stat_5 <- (p_hat_5 - p_0_5) / sqrt(p_0_5 * (1 - p_0_5) / sample_size_religion)

#one-tailed p-value
p_value_5 <- pnorm(z_t_stat_5)

#p-value
cat("One-tailed p-value:", p_value_5)
```

State the null hypothesis and interpret what the results tell us about the null and alternative hypotheses:

-   Null hypothesis = in countries where Muslim is the predominate religious group, TFR/women, have no difference in number of children they have on average.

-   Alternative: Countries where Muslim is the predominate religion, women have a higher TFR

<!-- -->

-   One tailed p-value is = 4.6775e-36

-   This is a very small value, providing evidence against the null hypothesis. The p-value being so small, this indicates that it's very unlikely to observe this data if there were no difference in the average number of children women have in countries where Muslim is the predominant religion vs. other countries w/ other predominate religious groups.

-   Based on the data presented, I reject the null hypothesis = no difference in average number of TFR

-   Results strongly support alternative hypothesis = TFR is higher on average where Muslim is the predominate group.
